A Fredholm operator approach to eigenvalue problems in elasticity and planar Stokes flow

### 2019-01-14 doc/draft-01/02prelim.tex
in which our hero updates the intro and adds some references
Updating paper
update text, add proof of radiation condition
update appendix
updating appendix
fix sign in def of bi-helmholtz green's function
update equation names

In this paper, vector-valued
quantities are denoted by bold, lower-case letters
(e.g. $\bh$), while tensor-valued quantities are bold
and upper-case (e.g. $\mathbf{T}$). 
Subscript indices of non-bold characters (e.g. $h_j$ or $T_{jkl}$)
are used to denote the entries within a vector ($\bh$) or tensor ($\bT$).
We use the standard Einstein summation convention; in other words, 
there is an implied sum taken over the repeated indices of 
any term (e.g. the symbol $a_{j} b_{j}$ is used to represent the sum
$\sum_{j} a_{j} b_{j}$).
If $\xx = (x_1,x_2)^\intercal$, then $\xx^\bot = (-x_2,x_1)^\intercal$.
Similarly, $\nabla^\bot = (-\partial_{x_2},\partial_{x_1})^\intercal$.

@@

Let $L_x$ denote a linear differential operator. A fundamental
solution $G(\xx,\yy)$ of $L_x$ satisfies the equation
$L_x G(\xx,\yy) = \delta_y(\xx)$ in the distributional sense, i.e.
for sufficiently smooth $f$
\begin{equation}
  L_x \int_{\R^2} G(\xx,\yy) f(\yy) \, dy = f(\xx) \; .
\end{equation}
We consider here
``free-space'' Green's functions, i.e. fundamental solutions which satisfy
some natural growth or decay conditions as $|\xx-\yy| \to \infty$.
The Green's function of the Helmholtz-biharmonic equation
\cref{eq:buck1} is given by 
\begin{equation}
  \Gbh(\xx,\yy) = -\frac{1}{k^2}
  \left (\frac{1}{2\pi} \log |\xx-\yy| +
  \frac{i}{4} H_0^{(1)}(k|\xx-\yy|) \right ) \, , 
\end{equation}
where $k$ is the Helmholtz paramater in the Helmholtz-Biharmonic equation,
and $H_{0}^{1}(r)$ is the Hankel function of the first kind of order zero.
Note that this is a scaled difference of the Green's function for
Laplace, i.e.

\begin{equation}
  \Glap(\xx,\yy) = -\frac{1}{2\pi} \log |\xx-\yy| \; ,
\end{equation}
and the Green's function for the Helmholtz equation

\begin{equation}
  \Ghelm(\xx,\yy) = \frac{i}{4} H_0^{(1)}(k|\xx-\yy|) \; .
\end{equation}

@@

In \cite{jiang2013second}, Jiang extended the Farkas formulation for the solution of the modified
biharmonic equation, albeit for different boundary conditions. With minor modifications, their formulation
can also be adopted for the solution of the Dirichlet problem for the Helmholtz-biharmonic equation
\begin{align}
\Delta (\Delta + k^2) w &= 0 \quad \text{in} \quad \Omega \label{eq:helmbi} \\
w &= f \quad \text{on} \quad \Gamma \\
\dwdn &=g \quad \text{on} \quad \Gamma \,,
\end{align}
where $f,g$ are prescribed functions on the boundary.

@@

\begin{lem}
Suppose that $\Omega \subset \R^2$ is a simply connected region whose boundary $\Gamma$ is $\cC^{2}$. 
Furthermore, suppose that $k$ is not a buckling eigenfrequency of $\Omega$, i.e., the only solution $w$ to~\cref{eq:buck1,eq:buck2,eq:buck3} is  $w \equiv 0$. Suppose further that $f,g \in \cC^{2}(\Gamma)$. 
Then there exists a unique solution $\sigma_{1},\sigma_{2} \in \cC^{2}$ which satisfies~\cref{eq:farkIntEq}. 
Furthermore, $w(\bx)$ defined via~\cref{eq:farkRep} satisfies the Helmholtz-Biharmonic equation~\cref{eq:helmbi}
along with the boundary conditions $w(\bx)=f(\bx)$ and $\dwdn(\bx) = g(\bx)$ for $\bx \in\Gamma$. 
\end{lem}

@@

The Stokes eigenvalue problem (with ``no-slip'' boundary conditions)
seeks values $k^2$ such that the equations
\begin{align}
  \nabla p - \Delta \uu - k^2 \uu &= 0 \quad \textrm{in} \quad \Omega
  \; , \label{eq:helmstokes}\\
  \nabla \cdot \uu &= 0 \quad \textrm{in} \quad \Omega
  \; , \label{eq:helmstokescty} \\
  \uu &= 0 \quad \textrm{on} \quad \Gamma \; \label{eq:helmstokesnoslip}
\end{align}
have a non-trivial solution. In the case that $k=i\alpha$ for
some real-valued $\alpha$, the equations \cref{eq:helmstokes,eq:helmstokescty}
are known as the modified Stokes equations and are of particular
interest for their application to numerical simulations of unsteady
flow
\cite{pozrikidis1992boundary,biros2002embedded,jiang2013second}.
In the absence of a preferred name for the case of real-valued $k$,
we will refer to \cref{eq:helmstokes,eq:helmstokescty} as the
Helmholtz-Stokes equations.

@@

\begin{equation}
  \GG = - \II \Delta \Gbh + \nabla \otimes \nabla \Gbh
\end{equation}
is then the equivalent of a Stokeslet
\cite{pozrikidis1992boundary} for the problem
\eqref{eq:helmstokes}.

@@

\begin{equation}
  \DD \bmu (\xx) = \int_\Gamma \left ( \TT_{\cdot,\cdot,k}\nu_k(\yy)
  \right )^\intercal \bmu(\yy) \, dS(\yy) \; ,
\end{equation}
where $\bnu$ denotes the outward unit normal to the boundary.
If we write $\bmu = \bnu \mu_\nu + \btau \mu_\tau$,
where $\btau = \bmu^\bot$ is the positively oriented unit
tangent to the curve, then we have

\begin{align}
  \left ( \TT_{\cdot,\cdot,k}\nu_k(\yy) \right )^\intercal
  \bmu(\yy) &= \left ( - \nabla \Glap(\xx,\yy) + 2 \nabla^\bot
  \partial_{\nu\tau} \Gbh(\xx,\yy) \right ) \mu_\nu(\yy) \nonumber \\
  & \qquad+
  \nabla^\bot \left (\partial_{\tau\tau}-\partial_{\nu\nu} \right )
  \Gbh(\xx,\yy) \mu_\tau(\yy) \; .
\end{align}

>>>>

In this paper, vector-valued
quantities are denoted by bold, lower-case letters
(e.g. $\bh$), while tensor-valued quantities are bold
and upper-case (e.g. $\mathbf{T}$). 
Subscript indices of non-bold characters (e.g. $h_j$ or $T_{jkl}$)
are used to denote the entries within a vector ($\bh$) or tensor ($\bT$).
We use the standard Einstein summation convention; in other words, 
there is an implied sum taken over the repeated indices of 
any term (e.g. the symbol $a_{j} b_{j}$ is used to represent the sum
$\sum_{j} a_{j} b_{j}$).
If $\xx = (x_1,x_2)^\intercal$, then $\xx^\bot = (-x_2,x_1)^\intercal$.
Similarly, $\nabla^\bot = (-\partial_{x_2},\partial_{x_1})^\intercal$.
Upper-case script characters (e.g. $\mathcal{K}$) are reserved for
operators on Banach spaces, with $\mathcal{I}$ denoting the
identity.

@@

Let $L_x$ denote a linear differential operator. A fundamental
solution $G(\xx,\yy)$ of $L_x$ satisfies the equation
$L_x G(\xx,\yy) = \delta_y(\xx)$ in the distributional sense, i.e.
for sufficiently smooth $f$
\begin{equation}
  L_x \int_{\R^2} G(\xx,\yy) f(\yy) \, dy = f(\xx) \; .
\end{equation}
We consider here
``free-space'' Green's functions, i.e. fundamental solutions which satisfy
some natural growth or decay conditions as $|\xx-\yy| \to \infty$.
The Green's function of the oscillatory biharmonic equation
\cref{eq:buck1} is given by 
\begin{equation}
  \Gbh(\xx,\yy) = \frac{1}{k^2}
  \left (\frac{1}{2\pi} \log |\xx-\yy| +
  \frac{i}{4} H_0^{(1)}(k|\xx-\yy|) \right ) \, , 
\end{equation}
where $k$ is the Helmholtz paramater in the oscillatory biharmonic equation,
and $H_{0}^{1}(r)$ is the Hankel function of the first kind of order zero.
Note that this is a scaled difference of the Green's function for
Laplace, i.e.

\begin{equation}
  \Glap(\xx,\yy) = \frac{1}{2\pi} \log |\xx-\yy| \; ,
\end{equation}
and the Green's function for the Helmholtz equation

\begin{equation}
  \Ghelm(\xx,\yy) = -\frac{i}{4} H_0^{(1)}(k|\xx-\yy|) \; .
\end{equation}

@@

In \cite{jiang2013second}, Jiang extended the Farkas formulation for the solution of the modified
biharmonic equation, albeit for different boundary conditions. With minor modifications, their formulation
can also be adopted for the solution of the Dirichlet problem for the oscillatory biharmonic equation
\begin{align}
\Delta (\Delta + k^2) w &= 0 \quad \text{in} \quad \Omega \label{eq:helmbi} \\
w &= f \quad \text{on} \quad \Gamma \label{eq:helmbibc1} \\
\dwdn &=g \quad \text{on} \quad \Gamma \label{eq:helmbibc2} \,,
\end{align}
where $f,g$ are prescribed functions on the boundary.

@@

\begin{lem}
Suppose that $\Omega \subset \R^2$ is a simply connected region whose boundary $\Gamma$ is $\cC^{2}$. 
Furthermore, suppose that $k$ is not a buckling eigenfrequency of $\Omega$, i.e., the only solution $w$ to~\cref{eq:buck1,eq:buck2,eq:buck3} is  $w \equiv 0$. Suppose further that $f,g \in \cC^{2}(\Gamma)$. 
Then there exists a unique solution $\sigma_{1},\sigma_{2} \in \cC^{2}$ which satisfies~\cref{eq:farkIntEq}. 
Furthermore, $w(\bx)$ defined via~\cref{eq:farkRep} satisfies the oscillatory biharmonic equation~\cref{eq:helmbi}
along with the boundary conditions $w(\bx)=f(\bx)$ and $\dwdn(\bx) = g(\bx)$ for $\bx \in\Gamma$.
\label{lem:farkinv}
\end{lem}

@@

The Stokes eigenvalue problem (with ``no-slip'' boundary conditions)
seeks values $k^2$ such that the equations
\begin{align}
  \nabla p - \Delta \uu - k^2 \uu &= 0 \quad \textrm{in} \quad \Omega
  \; , \label{eq:helmstokes}\\
  \nabla \cdot \uu &= 0 \quad \textrm{in} \quad \Omega
  \; , \label{eq:helmstokescty} \\
  \uu &= 0 \quad \textrm{on} \quad \Gamma \; \label{eq:helmstokesnoslip}
\end{align}
have a non-trivial solution. In the case that $k=i\alpha$ for
some real-valued $\alpha$, the equations \cref{eq:helmstokes,eq:helmstokescty}
are known as the modified Stokes equations and are of particular
interest for their application to numerical simulations of unsteady
flow
\cite{pozrikidis1992boundary,biros2002embedded,jiang2013second}.
In the absence of a preferred name for the case of real-valued $k$,
we will refer to \cref{eq:helmstokes,eq:helmstokescty} as the
oscillatory Stokes equations.

@@

\begin{equation} \label{eq:stokeslet}
  \GG = - \II \Delta \Gbh + \nabla \otimes \nabla \Gbh
\end{equation}
is then the equivalent of a Stokeslet
\cite{pozrikidis1992boundary} for the problem
\eqref{eq:helmstokes}.

@@

\begin{equation} \label{eq:doublelayer}
  \DD \bmu (\xx) = \int_\Gamma \left ( \TT_{\cdot,\cdot,k}\nu_k(\yy)
  \right )^\intercal \bmu(\yy) \, dS(\yy) \; ,
\end{equation}
where $\bnu$ denotes the outward unit normal to the boundary.
If we write $\bmu = \bnu \mu_\nu + \btau \mu_\tau$,
where $\btau = \bmu^\bot$ is the positively oriented unit
tangent to the curve, then we have

\begin{align} \label{eq:stokesdlkernel}
  \left ( \TT_{\cdot,\cdot,k}\nu_k(\yy) \right )^\intercal
  \bmu(\yy) &= \left ( - \nabla \Glap(\xx,\yy) + 2 \nabla^\bot
  \partial_{\nu\tau} \Gbh(\xx,\yy) \right ) \mu_\nu(\yy) \nonumber \\
  & \qquad+
  \nabla^\bot \left (\partial_{\tau\tau}-\partial_{\nu\nu} \right )
  \Gbh(\xx,\yy) \mu_\tau(\yy) \; .
\end{align}
